Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-05 paper/main.tex
Updates

Here, we have the option to choose the estimators based on the properties
(distribution, type of relationship, and so on) of the data. Non-parametric
ensemble estimators such as Random Forest and XGBoost can handle any
combination of data types while performing reasonably well on a wide range of
distributions and relationship types. For linear relationships, we can also
choose simpler models such as linear regression.

Repeating the above residualization method for both $ X $ and  $ Y $, we obtain
residual matrices $ R_X $ and $ R_Y $. The type of the variable determines the
shape for these matrices. If the variable is continuous or ordinal, the
residual matrix is of shape $ ( n \times 1 ) $ and if the variable is
categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where $ k
$ is the number of categories of the variable.

@@

We use Pillai's Trace for our purpose for two main reasons: 1) It uses all the
canonical correlation values, 2) Its interpretation is similar to correlation
coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
linear relationship. This gives us the following partial measure of
association:

@@

\begin{enumerate}
	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ (no
		association) to $ 1 $ (perfect linear association). This bounded
		nature simplifies the interpretation and allows for easy 
		comparisons across multiple values.
	\item \textbf{Independent of Sample Size: } The measure remains consistent
		regardless of the sample size ensuring reliability across 
		studies with different sample sizes.
	\item \textbf{Invariant to the Dummy Encoding: } We use dummy encoding 
		to compute the residual matrix for categorical variables. Since
		canonical correlations identify linear combinations across
		columns, this measure is invariant to the specific dummy
		encoding scheme used.
	\item \textbf{Equivalent to Partial Correlation}
		When both $ X $ and $ Y $ are continuous variables, our measure
		is equivalent to the absolute value of partial correlation.
	\item \textbf{Equivalent to polychoric and polyserial correlation for
			ordinal and continuous variables: }
		Under the assumption that the ordinal variable is generated by
		discretizing an underlying gaussian continuous variable, this
		effect size is equivalent to polychoric and polyserial
		correlation. As both of them recover the Pearson correlation
		coefficient.
	\item \textbf{Related to Cram\'er's V: } 
\end{enumerate}

>>>>

Here, we have the option to choose the estimators based on the properties
(distribution, type of relationship, and so on) of the data. Non-parametric
ensemble estimators such as Random Forest and XGBoost can handle any
combination of data types while performing reasonably well on a wide range of
distributions and relationship types. For linear relationships, we can also
choose simpler models such as linear regression and its variants.

Repeating the above residualization process for both $ X $ and  $ Y $, we
obtain residual matrices $ R_X $ and $ R_Y $. The type of the variable
determines the shape for these matrices. If the variable is continuous or
ordinal, the residual matrix is of shape $ ( n \times 1 ) $ and if the variable
is categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where
$ k $ is the number of categories of the variable.

@@

We use a slightly modified version of Pillai's Trace for our purpose for two
main reasons: 1) It uses all the canonical correlation values, 2) Its
interpretation is similar to correlation coefficient, i.e., $ 0 $ signifies no
association and $ 1 $ signifies perfect linear relationship. This gives us the
following partial measure of association:

@@

\begin{enumerate}
	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ (no
		association) to $ 1 $ (perfect linear association). This bounded
		nature simplifies the interpretation and allows for easy 
		comparisons across multiple values.
	\item \textbf{Independent of Sample Size: } 
	\item \textbf{Invariant to the Dummy Encoding: } We use dummy encoding 
		to compute the residual matrix for categorical variables. Since
		canonical correlations identify linear combinations across
		columns, this measure is invariant to the specific dummy
		encoding scheme used.
	\item \textbf{Equivalent to Partial Correlation}
		When both $ X $ and $ Y $ are continuous variables, our measure
		is equivalent to the absolute value of partial correlation.
	\item \textbf{Equivalent to polychoric and polyserial correlation for
			ordinal and continuous variables: }
		Under the assumption that the ordinal variable is generated by
		discretizing an underlying gaussian continuous variable, this
		effect size is equivalent to polychoric and polyserial
		correlation. As both of them recover the Pearson correlation
		coefficient.
	\item \textbf{Related to Cram\'er's V: } 
\end{enumerate}
